---
author: "Brendan Connors"
title: "Example of simulating stock-recruitment data, fitting a Dynamic Linear Model to it and summarizing estimation performance"
---
  
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
source("dlm-wrapper.R")
```
This is a very quick and dirty example of fitting a DLM to a bunch of simulated data across a range of stock sizes/depletion and then visualizing mean % bias followed by an example of creating a confusion matrix that quantifies type 1 and 2 error.

First let's simulate some stock-recruitment data assuming a Ricker type relationship with a linear declining trend in productivity and no change in capacity over time, and lets do it 100 times, for a range of stock sizes (implemented via a range of fixed exploitation rates):
  
```{r}

# source a ugly quick and dirty set of sims
source("ugly-toy-sims-depletion-scenarios.R")

head(srSimsdfERs)
```

Then lets fit a DLM, with only time varying alpha, to each Monte Carlo trial in each scenario:
  
```{r, warning=FALSE}
DLMOut<-srSimsdfERs
DLMOut<-cbind(srSimsdfERs,as.data.frame(matrix(NA,dim(srSimsdfERs)[1],4)))
  
colnames(DLMOut)<-c("scen","MC","byr","spwn","rec","alpha_true","beta_true", "alpha", "beta","alpha_se", 'beta_se')

for(j in unique(DLMOut$scen)){
  for(i in 1:nMC){
    dlm_model <- fitDLM(data = subset(DLMOut, scen==j & MC==i),
                        alpha_vary = TRUE,
                        beta_vary = TRUE)
    DLMOut[which(DLMOut$scen==j & DLMOut$MC==i),c(8:11)] <- dlm_model$results[,c(12:15)]
  }
}
```

Next calculate mean percent bias in alpha and beta: 
```{r}

pms <- DLMOut %>%
  group_by(scen,MC) %>%
  dplyr::summarize(
    alpha_mpb=mean((alpha_true-alpha)/alpha_true)*100,
    beta_mpb=mean((beta_true-beta)/beta_true)*100) %>%
  pivot_longer(alpha_mpb:beta_mpb,names_to="parameter",values_to="mpb")


```

And finally plot:

```{r}
ggplot(data=pms,aes(x=factor(scen), y=mpb, fill=parameter))+
  geom_boxplot(outlier.shape = NA)+
  xlab("Stock depletion") +
  ylab("Mean % bias") +
  coord_cartesian(ylim=c(-100,400))+
  scale_fill_viridis_d(labels=c("alpha", "beta"))+
  theme_bw() 

```

Now lets fit 4 different DLMS and figure out the type 1 and 2 errors based on AICc. First run sims:

```{r}

# source a ugly quick and dirty set of sims
source("ugly-toy-sims-confusion-scenarios.R")

head(srSimsdfConf)
```
Next fit four different DLMs to the simulated data so that we can estimate error rates, in this case for a stock depletion of 50% (i.e., fixed 50% harvest rate):  
```{r, warning=FALSE}
DLMOut<-srSimsdfConf

confusion<-matrix(0,400,6)
confusion<-as.data.frame(confusion)
confusion[,1]<-rep(unique(DLMOut$scen),each=100)
confusion[,2]<-rep(1:100,times=4)
colnames(confusion)<-c("scenario","MC","stationary","alpha","beta","alphabeta")

for(j in unique(DLMOut$scen)){
  for(i in 1:nMC){
    dlm_model_stat <- fitDLM(data = subset(DLMOut, scen==j & MC==i),
                        alpha_vary = FALSE,
                        beta_vary = FALSE)
    
    dlm_model_alpha <- fitDLM(data = subset(DLMOut, scen==j & MC==i),
                        alpha_vary = TRUE,
                        beta_vary = FALSE)
    
    dlm_model_beta <- fitDLM(data = subset(DLMOut, scen==j & MC==i),
                        alpha_vary = FALSE,
                        beta_vary = TRUE)
    
    dlm_model_both <- fitDLM(data = subset(DLMOut, scen==j & MC==i),
                        alpha_vary = TRUE,
                        beta_vary = TRUE)
    best_mod <- which.min(c(dlm_model_stat$AICc,dlm_model_alpha$AICc,dlm_model_beta$AICc,dlm_model_both$AICc))

    confusion[which(confusion$scen==j & confusion$MC==i),2+best_mod]<-1

      }
}

```

Then calculate proportion of trials that correctly and incorrectly true underlying dynamics, in this case based on model with lowest AICc:

```{r}
confusionLong <- confusion %>%
  group_by(scenario) %>%
  dplyr::summarize(
    pStation=sum(stationary)/100,
    pAlpha=sum(alpha)/100,
    pBeta=sum(beta)/100,
    pAlphaBeta=sum(alphabeta)/100) %>%
  pivot_longer(pStation:pAlphaBeta,names_to="estimated",values_to="pMC") %>%
  mutate(trueNum=case_when(scenario=="stationary"~1,
                           scenario=="downAlpha"~2,
                           scenario=="downBeta"~3,
                           scenario=="downAlphaBeta"~4)) %>%
  mutate(estNum=case_when(estimated=="pStation"~1,
                           estimated=="pAlpha"~2,
                           estimated=="pBeta"~3,
                           estimated=="pAlphaBeta"~4))  
```

And lastly plot confusion matrix:

```{r}
ggplot(data=confusionLong,aes(x=trueNum, y=estNum, fill=pMC))+
   geom_tile(aes(fill = pMC))+
  scale_fill_viridis_c(limits=c(0,1), name="Proportion\n of trials")+
  xlab("True") +
  ylab("Estiamted") +
  geom_text(aes(label = round(pMC, 2)), size = 4, color="white")+
  scale_x_continuous(labels=c("Stationary", "Alpha vary", "Beta vary", "Both vary"), breaks=c(1,2,3,4))+
  scale_y_continuous(labels=c("Stationary", "Alpha vary", "Beta vary", "Both vary"), breaks=c(1,2,3,4))+
  theme_bw() 
```
