---
author: "Brendan Connors"
title: "Example of simulating stock-recruitment data and then fitting a Dynamic Linear Model to it"
---

# Setup

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
source("dlm-wrapper.R")
```

# Simulate some stock-recruitment data

First we will simulate some stock-recruitment data assuming a Ricker type realtionship with a linear declining trend in productivity and no change in capacity over time:

```{r}
alpha <- seq(2,0.5,length.out=50) # Ricker a parameter
beta <- seq(0.01,0.01,length.out=50) # Ricker b parameter
seq <- alpha/beta # equilibrium population size

spwn <- rep(NA,50)
rec <- spwn
spwn[1] <- seq[1]*0.5

for(i in 1:49){
  rec[i] <- spwn[i]*exp(alpha[i]-beta[i]*spwn[i])*exp(rnorm(1,0,0.25))
  spwn[i+1] <-rec[i]*0.8
}

rec<-rec[1:49]
spwn<-spwn[1:49]
byr<-seq(1:49)

bt<-as.data.frame(cbind(byr,spwn,rec, alpha[-50], beta[-50]))
colnames(bt)<-c("byr","spwn","rec","alpha_true","beta_true")
```

Then bind everything together and create a data-frame that will play nice with DLM wrapper functions:

```{r}
bt<-as.data.frame(cbind(byr,spwn,rec, alpha[-50], beta[-50]))
colnames(bt)<-c("byr","spwn","rec","alpha_true","beta_true")
```

Next fit a Dynamic Linear Model to simulated data, in this case specifying the intercept (alpha) as time varying:

```{r, warning=FALSE}
dlm_model <- fitDLM(data = bt,
              alpha_vary = TRUE,
              beta_vary = FALSE)
```

And plot simulated stock-recruitment data and and compare true and estimated parameter estimates:

```{r, fig.width=8, fig.height=4}
plotDLM(dlm_model)
```
